An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullets

### 2021-09-02 CMPSpaper_writing/CMPSpaper_writing.Rmd
edits through results section

## Methods Maybe?

@@

$\mathrm{\overline{CMPS}_{diff}}$ highlights the difference between CMPS scores of matching and non-matching comparisons. If two bullets are non-matching, all 36 CMPS scores are expected to be small with relatively the same values, resulting in a $\mathrm{\overline{CMPS}_{diff}}$ value close to 0. For the example above, $\mathrm{\overline{CMPS}_{diff}} = 12.5 - 1.53 = 10.97$

**The Scaled CMPS Score**

Another issue with the CMPS score is that the highest possible CMPS score (the total number of basis segments) might differ across comparisons (as shown in \autoref{fig:tiles2}(a)) due to different lengths of bullet signatures and different lengths of basis segments specified by the parameter. A CMPS score of 5 might indicate a non-match if the highest possible CMPS score is 30 but indicate a match if the highest possible CMPS score is 6. Thus, we introduce the scaled CMPS score, denoted as $c^*_{ij}$. Let $s_{ij}$ denote the highest possible CMPS score or the total number of basis segments, we have

@@

The scaled CMPS scores of the above example are shown in \autoref{fig:tiles2}(b). \wj{Compared with the original CMPS scores, the scaled CMPS scores are more comparable in the sense that they all lie within the interval $(0, 1)$ no matter the choice of the length of the basis segments. Note that the scaled CMPS scores can be combined with the CMPS metrics introduced above to obtain their scaled version.} For example,  $\mathrm{\overline{CMPS^*}_{diff}} = 0.498$ for \autoref{fig:tiles2}(b)

@@

We now introduce a measure called "sum of squares ratio" that quantifies the separation between two groups. Let $n_T$ denote the total number of observations, $n_k$ denote the number of observations in group $k$, and $y_{kl}$ denote the $l$-th observation in group $k$, for $k = 1,2$ and $l = 1, \dots, n_k$. Let $\bar{y}_{k.} = \frac{1}{n_k} \sum_{l=1}^{n_k} y_{kl}$ denote the mean value in group $k$ and $\bar{y}_{..} = \frac{1}{n_T} \left(  \sum_{k} \sum_{l = 1}^{n_k} y_{kl} \right)$ denote the mean value of all observations. Since we are interested in the separation between two groups, we can model $y_{kl}$ as:

@@

\begin{align}
V = \frac{\sum_k n_k (\bar{y}_{k.} - \bar{y}_{..})^2}{\sum_k \sum_l^{n_k} (y_{kl} - \bar{y}_{k.})^2 }
\end{align}

@@

\wj{add citation for one way ANOVA?}

With the sum of squares ratio $V$ being defined, we are able to construct a pipeline that aims to find the optimal parameter value for the CMPS algorithm maximizing the sum of squares ratio. In the following section, we will use the sum of squares ratio to compare the CMPS metrics introduced above and investigate effects of different sets of parameters.

\wj{we might not find the real optimal parameter since the parameter space of npeaks.set is too complicated...}

@@

As presented in the work of @cmps, researchers applied the CMPS method to one of the Hamby set scans (Hamby 252 \wj{to the best of our knowledge} ). In order to show that our implementation of the CMPS algorithm is able to reproduce the results in @cmps and be used for other data sets, we applied our implementation to both Hamby set 252 and Hamby set 44. \wj{For the purpose of reproducibility, we present how we obtained bullet signatures from the Hamby set data and include code in the appendix}: for both Hamby 252 and Hamby 44, we started with x3p files in the database, followed the framework proposed by @aoas with the same set of parameters, removed damaged bullet scans, obtained bullet signatures for each bullet land engraving, and removed outliers in bullet signatures. Note that researchers of @cmps \wj{applied the CMPS algorithm to bullet signatures as well but used a framework different from ours.} However, since their work is not open-source, we were not able to follow their framework and were only able to reproduce the results for Hamby set 252 qualitatively.

@@

As noted above, the CMPS scores we found here are not exactly the same as those presented in @cmps since we were not able to follow their framework, but the results presented in \autoref{fig:result1_252} are qualitatively equivalent to those presented in @cmps, showing a clear separation between the known matching comparisons (KM) group and the known non-matching comparisons (KNM) group for both $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$.

Additionally, to mimic the parameters used in @cmps, we set `seg_length = 50` and `Tx = 25` to make sure that each basis segment has length of 78.125 \textmu m and the tolerance zone is $\pm 39.0625$ \textmu m (one unit represents 1.5625 \textmu for Hamby set 252).

The sum of squares ratios are 20.85 and 29.32 for $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively. \wj{TODO: add some comments following this result}

>>>>

## Evaluation Metrics

@@

$\mathrm{\overline{CMPS}_{diff}}$ highlights the difference between CMPS scores of matching and non-matching comparisons. If two bullets are non-matching, all 36 CMPS scores are expected to be small with relatively the same values, resulting in a $\mathrm{\overline{CMPS}_{diff}}$ value close to 0. For the example above, $\mathrm{\overline{CMPS}_{diff}} = 12.5 - 1.53 = 10.97$ <br>

**Scaled CMPS Scores**

Another issue with the CMPS score is that the highest possible CMPS score (the total number of basis segments) might differ across comparisons (as shown in \autoref{fig:tiles2}(a)) due to different lengths of bullet signatures and different lengths of basis segments specified by the parameter. A CMPS score of 5 might indicate a non-match if the highest possible CMPS score is 30 but indicate a match if the highest possible CMPS score is 6. Thus, we introduce the scaled CMPS score, denoted as $c^*_{ij}$. Let $s_{ij}$ denote the highest possible CMPS score or the total number of basis segments, then the scaled CMPS score $c^*_{ij}$ is defined as the ratio of raw score and maximum score:

@@

The scaled CMPS scores of the above example are shown in \autoref{fig:tiles2}(b). Compared to the original CMPS scores, scaled scores have values within the interval $(0, 1)$ regardless of the length of the basis segments and therefore make a comparison of values possible across different parameter choices. Similarly to the original CMPS scores, we will denote the scaled CMPS scores adjusted for out-of-phase background values by $\mathrm{\overline{CMPS^*}_{diff}}$.
For example,  $\mathrm{\overline{CMPS^*}_{diff}} = 0.498$ for \autoref{fig:tiles2}(b)

@@

The "sum of squares ratio"  quantifies how well two groups of values separate.  Let $n_T$ denote the total number of observations, $n_k$ denote the number of observations in group $k$, and $y_{kl}$ denote the $l$-th observation in group $k$, for $k = 1,2$ and $l = 1, \dots, n_k$. Let $\bar{y}_{k.} = \frac{1}{n_k} \sum_{l=1}^{n_k} y_{kl}$ denote the mean value in group $k$ and $\bar{y}_{..} = \frac{1}{n_T} \left(  \sum_{k} \sum_{l = 1}^{n_k} y_{kl} \right)$ denote the mean value of all observations. 
Consider the following model:

@@

\begin{align}
V = \frac{\sum_k n_k (\bar{y}_{k.} - \bar{y}_{..})^2}{\sum_k \sum_l^{n_k} (y_{kl} - \bar{y}_{k.})^2 }.
\end{align}

@@

Using the sum of squares ratio $V$ as an evaluation metric, we are able to construct a pipeline that aims to find the optimal parameter value for the CMPS algorithm by maximizing the sum of squares ratio. In the following section, we will use this ratio to compare the CMPS metrics introduced earlier and investigate effects of different sets of parameters.

@@

As presented in the work of @cmps, researchers applied the CMPS method to scans of one of the Hamby sets. While it is not explicitly stated in the paper, we presume this to be Hamby 252, as only those scans were publicly available at the time. In order to show that our implementation of the CMPS algorithm is able to reproduce the results in @cmps and be used for other data sets, we applied our implementation to both Hamby set 252 and Hamby set 44. For the purpose of reproducibility, we present how we obtained bullet signatures from the Hamby set data and include code in the appendix: for both Hamby 252 and Hamby 44, we started with scans in the form of x3p files in the database. Following the framework proposed by @aoas, we used the same set of parameters, removed damaged bullet scans, obtained bullet signatures for each bullet land engraving, and removed outliers in bullet signatures. Note that researchers of @cmps applied the CMPS algorithm to bullet signatures as well but used a framework different from ours. However, since their work is not open-source, we were not able to follow their framework and were only able to reproduce the results for Hamby set 252 qualitatively.

@@

As noted above, the CMPS scores we found here are not exactly the same as those presented in @cmps since we were not able to follow their framework, but the results presented in \autoref{fig:result1_252} are qualitatively equivalent to those presented in @cmps, showing a clear separation between scores based on comparisons from known matches (KM)  and scores from comparisons of known non-matches (KNM)  for both $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$.

Additionally, to mimic the parameters used in @cmps, we set `seg_length = 50` and `Tx = 25` to make sure that each basis segment has length of 78.125 \textmu m and the tolerance zone is $\pm 39.0625$ \textmu m (one unit represents 1.5625 \textmu m for Hamby set 252).

The sum of squares ratios are 20.85 and 29.32 for $\mathrm{CMPS_{max}}$ and $\mathrm{\overline{CMPS}_{max}}$, respectively. \hh{This indicates, that even though scores from $\mathrm{CMPS_{max}}$ for known-match comparisons are larger than scores from the averaged version of $\mathrm{\overline{CMPS}_{max}}$, these scores achieve a better separation between the two groups of comparisons.}
